LOSD: A Linked Open Social Dataset for Scientific Research

### 2019-07-02 mdpi/template.tex
revisado supplementary document e passado spellcheck

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
The fields of social network analysis and complex networks
are widely researched.
In fact, a myriad of results have been reported which are based in
diverse data, most often not accessible to researchers other than the publishing authors.
In order to enhance the scientific community's common repertoire,
this work presents an open dataset with diverse provenance of 
human social online networking.
Current data was obtained in 2013-16 from well-known online social networking platforms:
Facebook, Twitter, IRC, Email.
Data from Participa.BR, AA, and Cidade Democrática are also supplied as samples of less known instances.
The linked data representation was adopted in order to
comply with current best practices, homogenize access, facilitate discovery
and analyzes which integrate third party and provided data,
and make available a dataset with multiple metadata that admits on-demand retrieval of structures.
Futhermore, the method for obtaing and making the data available is potentially novel,
and may be of use for collecting and sharing additional data and for other research groups.
This document presents a description, usage remarks and overall statistics of the dataset,
which should favor subsequent work in complex and social networks and enhancements to the dataset.
}

@@

\begin{document}
\section{Summary}
The research on human social and complex networks often
relies on data that express real phenomena~\cite{c1,c2,c3}.
Such data is regularly not made publicly available,
which hinders validation of the results
and does not favor subsequent work.
Furthermore, there is a shortcoming of open datasets for benchmarking results
related to complex and social networks,
yielding diverse results from poorly related sources.
Social networks linked datasets are currently most often restricted to a specific platform
or express statistics of social networks, as described in Section~\ref{srel}.
This work presents 
a dataset named LOSD (Linked Open Social Data)
that has been achieved by linking social data derived from diverse
provenance, namely: Facebook, Twitter, IRC, Email groups, Participa.BR, Cidade Democrática, and AA.
Such data may be of use as a common and public repertoire for scientific
research involving complex networks, specially if social networks and textual content are pinpointed.
In fact, LOSD has already been used in scientific research~\cite{stab,thesis} (CNPq grant 140860/2013-4),
and should be further exploited e.g. in researching complex networks and data visualization (FAPESP grant 2017/05838-3)
which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.
Data collection was performed by a potentially innovative method which places the researcher
at the core of data collection, and keeps the resources (e.g. code, data, documents) open,
in order to ammeliorate ethical issues involved in studying human social systems.

@@

\subsection{Related work}\label{srel}
Most available datasets on human social networks rely only on one source (e.g.~\cite{nat,fb1}),
on specialized communities such as related to research or movies (e.g.~\cite{s1}),
or present only statistics about social networks, and not the data that yields the networks (e.g.~\cite{st1}.
In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud,
and made available a dataset from the Advogato platform,
dedicated to free software developers connected by trust relatioships.
There are various datasets which provide data obtained from Twitter (e.g.~\cite{tw1,tw2}),
most often with a specific focus (users, retweets, hashtags, user mentions, etc)
and not presented to the scientific community by means of a research article.
The datasets which are most closely related to LOSD are: KONECT~\cite{konect},
SNAP Datasets~\cite{snapnets}, the Network Repository~\cite{nr}, and ICON~\cite{icon}.
None of them are (RDF) linked data but express the networks in formats specialized for networks.
In total, these datasets make available thousands of networks,
a fraction of them are social networks from diverse provenance and
of various types (e.g. simple, directed, weighted, bipartite, temporal networks).
Compared to LOSD, KONECT, SNAP, Network Repository, and ICON are complementary, providing a further diversity of networks,
but only as simplified data, i.e. in specialized formats, most often expressed as columns of node ids
that are connected and metadata (such as timestaps for temporal networks).
In summary, no linked human social network dataset presented to the
scientific community containts the diverse provenance used in LOSD,
or provides such a wide range of metadata,
potentially because researchers did not have the means to acquire and
share the data as described in Section~\ref{acq},
and because linked data standards are (still) not widely adopted in 
every research domain.

@@

\subsection{Snapshots}
Of central importance to the presented dataset is the concept of a snapshot.
A snapshot is herein a set of data gathered together, at a contiguous time
unit.
Examples: the first 20 thousand email messages of an email list
comprises a snapshot; the tweets from the MAMA music event is a
snapshot; the friendship, interaction and posts structures of a facebook
group, prospected at the same time, is a snapshot.
Table~\ref{tsnap} denotes the number of snapshots from each provenance in LOSD.

@@

\section{Methods}\label{methods}
LOSD is expressed as RDF linked data
and is ontologicaly described through the
data-driven ontology synthesis method presented in Section~\ref{ont}.

\subsection{Data acquisition}\label{acq}
Research involving human subjects raises well-known ethical issues,
although ethical regulation for (non-medical) research in social systems
is not always endorsed by scientists~\cite{eth}.
Furthermore, the online social networking platforms often restrict access
to the data and users may consider illegal the use of their data through web crowlers~\cite{ile1,ile2,ile3}.
Although Twitter and GMANE (email) data are available through public APIs,
the other sources are more problematic.
In this context, research using data, e.g. from Facebook, is precarious
if not prohibited.
We reached a potential solution to this issue through continuated research with
collaborators in the fields of phylosophy, social sciences, antropology and psychology.
The conceptual framework is presented in~\cite{antphy,antphy2,antphy3}, and in essence
consists in researching first the social networks related to the researchers,
and keeping the procedures, data, algorithms and results publicly available.
These practices enabled collecting and sharing the data in LOSD,
and was called ``Anthropological Physics'' because it relies in the
etnographic tecnique of investigating social structures
by studying the researcher, as to ammeliorate the ethical issues
(not to be confused e.g. with ``virtual ethnography''~\cite{veth}).
Accordingly, aside from the data acquired from Twitter, all data in LOSD
was gathered from social instances where the first author was a participant.
For example, all Facebook ego networks were donated to the first author by
users which knew him and were interested in his research and intervention on the networks,
and all email and Facebook groups data was obtained from groups where the first author was a member.

Facebook data was obtained through the Netvizz software~\cite{netviz},
when Facebook allowed downloading friendships and interaction data as available in LOSD.
Data from Twitter was obtained through the Twitter streamming API.
Data From IRC was obtained through bots that kept logs of the messages.
Email data was obtained through the GMANE public API.
Data from Participa.BR, Cidade Democrática, and AA was obtained through contact
of the researcher with the developers of the platforms.
These sources are expreesed in Table~\ref{tab:provenance}.

@@

One may also write an ontology using the OWL protocol,
with which all the expressive capabilities of RDFS are
available, but one can also, among other things~\cite{ont}:
\begin{itemize}
  \item state ``property axioms'', i.e. specify if a property is e.g. reflexive or transitive;
  \item state ``class restrictions'', i.e. specify if a class instance e.g. necessarily holds a relation to another class instance or data.
\end{itemize}
OWL has a richer vocabulary than RDFS and is (way more) complex.
This complexity is a drawback together with the greater computational cost
for performing inference.
The advantage is the greater power to represent conceptualizations.
Using ontologies enables automated reasoning (usually called ``inference'').
A very simple example: if a property \te{:r} is known to relate only mokeys to trees,
then, if there is a triple \te{:a :r :b}, the instance \te{:a} is considered a monkey,
and instance \te{:b} is considered a tree,
even if there is no explicit declaration similar
to \te{:a rdf:type :Mokey}.

@@

The authors chose to express all URIs in LOSD locally (i.e. within a local namespace), and linkage with external namespaces may be achieved e.g. by using subclass and ``same as'' relations.
Also, each snapshot holds a specific \te{po:Observation} instance related to each participant,
allowing for specific metadata such as provided by the original data or as may be provided uppon analysis.
Current LOSD does not hold statistics about each snapshot, or linkage to external namespaces (such as FOAF and Dublin Core),
to avoid overcomplicating this initial contribution and unecessary use of storage space.

@@

or on multigraphs (or muldimensional networks),

@@

  \item Overall facebook friendship network available in LOSD:

@@

  \item Overall facebook friendship network derived only from ego networks:

@@

  \item Overall facebook interaction network available in LOSD:

@@

  \item Email interaction network for email list shapshot with URI \verb|snapshot_uri|,
    with the all (preliminarily cleaned) text related to each message (node atribute):

@@

  \item Friendship network from Participa.BR, with the date when each frienship was established:

@@

This is not an exaustive list.
For a less canonical example for obtaining networks,
links may be derived from sufficiently similar vocabulary used by participants,
or from activity with sufficiently similar dates,
which yield networks which are not (standard) interaction or friendship networks.
Figure~\ref{fnet} describes most often criteria used to obtain social networks.
% mk figure with two participants and their relation though response, retweet, or friendship TTM

@@

\section{Conclusions and further work}
\label{conclusions}
This article outlined LOSD, a large dataset of human and complex networks, with almost 32 million triples and diverse provenance.
Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, were also described to the convenience of the newcommer
and for a consistent exposition.
All data is publicly available to provide the scientific community with a consistent and curated linked data dataset
that enables research validation, continuity, and benchmarking.
The diagrams and tables of the Supplementary document of this article
has further directions on the available structures and is a complementary overview.

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{
The fields of social network analysis and complex networks
are widely researched.
In fact, a myriad of results have been reported which are based in
diverse data, most often not accessible to researchers other than the publishing authors.
In order to enhance the scientific community's common repertoire,
this work presents an open dataset with diverse provenance of 
human social online networking.
Current data was obtained in 2013-16 from well-known online social networking platforms:
Facebook, Twitter, IRC, Email.
Data from Participa.BR, AA, and Cidade Democrática are also supplied as samples of less known instances.
The linked data representation was adopted in order to
comply with current best practices, homogenize access, facilitate discovery
and analyzes which integrate third party and provided data,
and make available a dataset with multiple metadata that admits on-demand retrieval of structures.
Furthermore, the method for obtaining and making the data available is potentially novel,
and may be of use for collecting and sharing additional data and for other research groups.
This document presents a description, usage remarks and overall statistics of the dataset,
which should favor subsequent work in complex and social networks and enhancements to the dataset.
}

@@

\begin{document}
\section{Summary}
The research on human social and complex networks often
relies on data that express real phenomena~\cite{c1,c2,c3}.
Such data is regularly not made publicly available,
which hinders validation of the results
and does not favor subsequent work.
Furthermore, there is a shortcoming of open datasets for benchmarking results
related to complex and social networks,
yielding diverse results from poorly related sources.
Social networks linked datasets are currently most often restricted to a specific platform
or express statistics of social networks, as described in Section~\ref{srel}.
This work presents 
a dataset named LOSD (Linked Open Social Data)
that has been achieved by linking social data derived from diverse
provenance, namely: Facebook, Twitter, IRC, Email groups, Participa.BR, Cidade Democrática, and AA.
Such data may be of use as a common and public repertoire for scientific
research involving complex networks, specially if social networks and textual content are pinpointed.
In fact, LOSD has already been used in scientific research~\cite{stab,thesis} (CNPq grant 140860/2013-4),
and should be further exploited e.g. in researching complex networks and data visualization (FAPESP grant 2017/05838-3)
which relies on networks enriched by textual data and on dynamic (or time-evolving) networks.
Data collection was performed by a potentially innovative method which places the researcher
at the core of data collection, and keeps the resources (e.g. code, data, documents) open,
in order to ameliorate ethical issues involved in studying human social systems.

@@

\subsection{Related work}\label{srel}
Most available datasets on human social networks rely only on one source (e.g.~\cite{nat,fb1}),
on specialized communities such as related to research or movies (e.g.~\cite{s1}),
or present only statistics about social networks, and not the data that yields the networks (e.g.~\cite{st1}.
In~\cite{foaf1}, the authors presented a set of tools to obtain linked social data available in the public data cloud,
and made available a dataset from the Advogato platform,
dedicated to free software developers connected by trust relationships.
There are various datasets which provide data obtained from Twitter (e.g.~\cite{tw1,tw2}),
most often with a specific focus (users, retweets, hashtags, user mentions, etc)
and not presented to the scientific community by means of a research article.
The datasets which are most closely related to LOSD are: KONECT~\cite{konect},
SNAP Datasets~\cite{snapnets}, the Network Repository~\cite{nr}, and ICON~\cite{icon}.
None of them are (RDF) linked data but express the networks in formats specialized for networks.
In total, these datasets make available thousands of networks,
a fraction of them are social networks from diverse provenance and
of various types (e.g. simple, directed, weighted, bipartite, temporal networks).
Compared to LOSD, KONECT, SNAP, Network Repository, and ICON are complementary, providing a further diversity of networks,
but only as simplified data, i.e. in specialized formats, most often expressed as columns of node ids
that are connected and metadata (such as timestamps for temporal networks).
In summary, no linked human social network dataset presented to the
scientific community contains the diverse provenance used in LOSD,
or provides such a wide range of metadata,
potentially because researchers did not have the means to acquire and
share the data as described in Section~\ref{acq},
and because linked data standards are (still) not widely adopted in 
every research domain.

@@

\subsection{Snapshots}
Of central importance to the presented dataset is the concept of a snapshot.
A snapshot is herein a set of data gathered together, at a contiguous time
unit.
Examples: the first 20 thousand email messages of an email list
comprises a snapshot; the tweets from the MAMA music event is a
snapshot; the friendship, interaction and posts structures of a Facebook
group, prospected at the same time, is a snapshot.
Table~\ref{tsnap} denotes the number of snapshots from each provenance in LOSD.

@@

\section{Methods}\label{methods}
LOSD is expressed as RDF linked data
and is ontologically described through the
data-driven ontology synthesis method presented in Section~\ref{ont}.

\subsection{Data acquisition}\label{acq}
Research involving human subjects raises well-known ethical issues,
although ethical regulation for (non-medical) research in social systems
is not always endorsed by scientists~\cite{eth}.
Furthermore, the online social networking platforms often restrict access
to the data and users may consider illegal the use of their data through web crowlers~\cite{ile1,ile2,ile3}.
Although Twitter and GMANE (email) data are available through public APIs,
the other sources are more problematic.
In this context, research using data, e.g. from Facebook, is precarious
if not prohibited.
We reached a potential solution to this issue through continued research with
collaborators in the fields of philosophy, social sciences, anthropology and psychology.
The conceptual framework is presented in~\cite{antphy,antphy2,antphy3}, and in essence
consists in researching first the social networks related to the researchers,
and keeping the procedures, data, algorithms and results publicly available.
These practices enabled collecting and sharing the data in LOSD,
and was called ``Anthropological Physics'' because it relies in the
ethnographic technique of investigating social structures
by studying the researcher, as to ameliorate the ethical issues
(not to be confused e.g. with ``virtual ethnography''~\cite{veth}).
Accordingly, aside from the data acquired from Twitter, all data in LOSD
was gathered from social instances where the first author was a participant.
For example, all Facebook ego networks were donated to the first author by
users which knew him and were interested in his research and intervention on the networks,
and all email and Facebook groups data was obtained from groups where the first author was a member.

Facebook data was obtained through the Netvizz software~\cite{netviz},
when Facebook allowed downloading friendships and interaction data as available in LOSD.
Data from Twitter was obtained through the Twitter streaming API.
Data From IRC was obtained through bots that kept logs of the messages.
Email data was obtained through the GMANE public API.
Data from Participa.BR, Cidade Democrática, and AA was obtained through contact
of the researcher with the developers of the platforms.
These sources are expressed in Table~\ref{tab:provenance}.

@@

One may also write an ontology using the OWL protocol,
with which all the expressive capabilities of RDFS are
available, but one can also, among other things~\cite{ont}:
\begin{itemize}
  \item state ``property axioms'', i.e. specify if a property is e.g. reflexive or transitive;
  \item state ``class restrictions'', i.e. specify if a class instance e.g. necessarily holds a relation to another class instance or data.
\end{itemize}
OWL has a richer vocabulary than RDFS and is (way more) complex.
This complexity is a drawback together with the greater computational cost
for performing inference.
The advantage is the greater power to represent conceptualizations.
Using ontologies enables automated reasoning (usually called ``inference'').
A very simple example: if a property \te{:r} is known to relate only monkeys to trees,
then, if there is a triple \te{:a :r :b}, the instance \te{:a} is considered a monkey,
and instance \te{:b} is considered a tree,
even if there is no explicit declaration similar
to \te{:a rdf:type :Monkey}.

@@

The authors chose to express all URIs in LOSD locally (i.e. within a local namespace), and linkage with external namespaces may be achieved e.g. by using subclass and ``same as'' relations.
Also, each snapshot holds a specific \te{po:Observation} instance related to each participant,
allowing for specific metadata such as provided by the original data or as may be provided upon analysis.
Current LOSD does not hold statistics about each snapshot, or linkage to external namespaces (such as FOAF and Dublin Core),
to avoid overcomplicating this initial contribution and unnecessary use of storage space.

@@

or on multigraphs (or multidimensional networks),

@@

  \item Overall Facebook friendship network available in LOSD:

@@

  \item Overall Facebook friendship network derived only from ego networks:

@@

  \item Overall Facebook interaction network available in LOSD:

@@

  \item Email interaction network for email list snapshot with URI \verb|snapshot_uri|,
    with the all (preliminarily cleaned) text related to each message (node attribute):

@@

  \item Friendship network from Participa.BR, with the date when each friendship was established:

@@

This is not an exhaustive list.
For a less canonical example for obtaining networks,
links may be derived from sufficiently similar vocabulary used by participants,
or from activity with sufficiently similar dates,
which yield networks which are not (standard) interaction or friendship networks.
Figure~\ref{fnet} describes most often criteria used to obtain social networks.

@@

\section{Conclusions and further work}
\label{conclusions}
This article outlined LOSD, a large dataset of human and complex networks, with almost 32 million triples and diverse provenance.
Tools and procedures used to achieve LOSD, and fundamental queries needed to retrieve the data, were also described to the convenience of the newcomer
and for a consistent exposition.
All data is publicly available to provide the scientific community with a consistent and curated linked data dataset
that enables research validation, continuity, and benchmarking.
The diagrams and tables of the Supplementary document of this article
has further directions on the available structures and is a complementary overview.
